% This thesis template is based on the personal template of Felix Moebius. Thanks, Felix!
% It has been extended and modified by Tobias Pfandzelter at the Scalable Software Systems group of TU Berlin.
% It is licensed under the terms of the MIT license, meaning you are free to use it however you see fit but we accept no liability.
% Good luck writing your thesis!

\documentclass[a4paper, 11pt]{article}

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fix-cm}

\usepackage[a4paper, margin=3cm]{geometry}
\usepackage[titletoc, title]{appendix}

\usepackage{color}
\usepackage{booktabs}
\usepackage[all]{nowidow}
\usepackage[dvipsnames]{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{acronym}
\usepackage{graphicx}
\usepackage{url}
\usepackage{titlesec}
\usepackage{csquotes}

\usepackage{transparent}
\usepackage{eso-pic}
\usepackage[section]{placeins}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{subcaption}
\usepackage{color,soul}

\renewcommand\thefigure{%
\thesection.\arabic{figure}}
\renewcommand\thesubfigure{%
\thesection.\arabic{figure}.\arabic{subfigure}}
\renewcommand\thetable{%
\thesection.\arabic{table}}

\usepackage[main=english, ngerman]{babel}

% we use the cleveref package to refer to figures, sections, etc.
% instead of "Figure~\ref{fig:example}", write only "\cref{fig:example}" and the word "Figure" (or table, etc) will be inserted normally
\usepackage[noabbrev,capitalise]{cleveref}

\usepackage[
    maxbibnames=99,
    style=alphabetic,
    url=false,
    backend=bibtex8,
    sortcites=true,
]{biblatex}
\addbibresource{refs.bib}
\DeclareFieldFormat[online]{urldate}{Last accessed: #1}
\DeclareFieldFormat{eprint}{arXiv: \href{https://arxiv.org/abs/#1}{#1}}
\DeclareFieldFormat[report]{title}{``#1''}


\newcommand{\projectTitle}{Exploring Datastore Access Latency across AWS Compute Services}
\newcommand{\thesisType}{Bachelor's Thesis}
\newcommand{\authors}{Bhupendra Dev Singh}
\newcommand{\matrikel}{464877}
\newcommand{\authorEmail}{\href{mailto:b.singh@campus.tu-berlin.de}{b.singh@campus.tu-berlin.de}}
\newcommand{\examinera}{Prof.~Dr.-Ing.~David Bermbach}
\newcommand{\examinerb}{Prof.~Dr.~habil.~Odej Kao}
\newcommand{\supervisor}{Trever Schirmer}

\newcommand{\projectYear}{2024}
\newcommand{\facultyName}{Fakultät Elektrotechnik und Informatik}
\newcommand{\departmentName}{Fachgebiet Scalable Software Systems}

\begin{document}
\input{front.tex}

\newpage

\section*{Abstract}
Cloud computing has become essential for modern applications, offering scalable and flexible services. Amazon Web Services (AWS) is a leading provider, yet little empirical data exists on how its compute and datastore services interact in terms of access latency. This thesis benchmarks access latency between EC2 and Lambda compute services paired with RDS, DynamoDB, and S3 datastores under constant and burst workloads. The results reveal that EC2-based pairs consistently deliver lower latency and more stable performance compared to Lambda-based pairs, which show greater variability. These insights can guide cloud architects in designing latency-sensitive applications. However, limitations such as simplified configurations, AWS Free Tier constraints, and single-region testing highlight the need for further studies to explore more complex and distributed scenarios. This thesis provides initial foundational insights that can serve as a basis for further exploration and development.

% GERMAN
\clearpage
\begin{otherlanguage}
    {ngerman}
    \section*{Kurzfassung}
    Cloud-Computing ist unverzichtbar für moderne Anwendungen geworden, da es skalierbare und flexible Dienste bietet. Amazon Web Services (AWS) ist ein führender Anbieter, jedoch es gibt nur wenige empirische Daten darüber, wie seine Compute- und Datenspeicherdienste in Bezug auf Zugriffslatenz zusammenwirken. Diese Arbeit benchmarkt die Zugriffslatenz zwischen den Compute-Diensten EC2 und Lambda in Kombination mit den Datenspeicherdiensten RDS, DynamoDB und S3 unter konstanten und variierenden Arbeitslasten. Die Ergebnisse zeigen, dass EC2-basierte Paare konsistent niedrigere Latenzen und eine stabilere Leistung liefern als Lambda-basierte Paare, die größere Schwankungen aufweisen. Diese Erkenntnisse können Cloud-Architekten dabei unterstützen, latenzsensible Anwendungen zu entwickeln. Einschränkungen wie vereinfachte Konfigurationen, die Beschränkungen des AWS Free Tiers und die Begrenzung auf eine einzige Region verdeutlichen jedoch die Notwendigkeit weiterer Studien, um komplexere und verteilte Szenarien zu untersuchen. Diese Arbeit liefert erste grundlegende Einblicke, die als Basis für weitere Untersuchungen und Entwicklungen dienen können.
\end{otherlanguage}

\newpage

\section*{Acknowledgements}
I would like to express my sincere gratitude to my supervisor, Trever Schirmer (Research Associate and PhD student at the Scalable Software Systems TU Berlin), for his invaluable guidance, constructive feedback, and continuous support throughout the development of this thesis. His expertise and encouragement have been instrumental in shaping my work and deepening my understanding of the subject matter.

I acknowledge the use of ChatGPT (GPT-4o and GPT-4o mini, OpenAI, https://chatgpt.com/) and Grammarly (https://grammarly.com/) for paraphrasing, language improvement, and text correction in this thesis.

\clearpage

\tableofcontents

\include{chapters/01_introduction.tex}
\clearpage
\include{chapters/02_background.tex}
\clearpage
\include{chapters/03_approach.tex}
\clearpage
\include{chapters/04_implementation.tex}
\clearpage
\include{chapters/05_results.tex}
\clearpage
\include{chapters/06_discussion.tex}
\clearpage
\include{chapters/07_relatedwork.tex}
\clearpage
\include{chapters/08_conclusion.tex}

\newpage
\printbibliography
\end{document}
